Cover letter for the original IJRR submission

### 2017-09-27 body/equilibrium.tex
typos
starting modifications

\subsection{A criterion for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in legged locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

We first give a linear program (LP) that verifies whether a contact configuration allows for static equilibrium.
From this formulation we derive a new LP that quantifies the robustness of the equilibrium to uncertainties in the contact forces.
In turn, from this value we can either choose the most robust candidate, or set a threshold on the required robustness. While the presented LP is original, it is based on an analysis of the problem that we proposed in \citep{Prete2016}, where the interested reader can find more details.

\subsubsection{Conditions for static equilibrium}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $m \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{\boldsymbol\upgamma}_{i1},\mathbf{\boldsymbol\upgamma}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{\boldsymbol\upgamma}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the nonslipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} -\mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i2} & \mathbf{n}_{i} - \mu \mathbf{\boldsymbol\upgamma}_{i2}}^T
\end{equation*}

Any force belonging to the linearized cone
can thus be expressed as a positive combination of its four generating rays.
%~ \deleted{Therefore we can express the nonslipping constraint on $\mathbf{f}_i$ as:}

@@

%~ \begin{align}
%~ \underbrace{\mat{m (\ddot{\mathbf{c}} - \mathbf{g})  \\ m \mathbf{c} \times (\ddot{\mathbf{c}} - \mathbf{g}) + \dot{\mathbf{L}}}}_\mathbf{w}
%~ = 
%~ \underbrace{
%~ \mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
%~ \hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
%~ }_\mathbf{G}
%~ \bm{\beta},
%~ \end{align}
%~ where $\mathbf{w}\in \Rv{6}$ is the so-called gravito-inertial wrench (GIW) \citep{qiu:dhm:2011, Caron2015} and $\hat{\mathbf{p}} \in \R{3}{3}$ is the cross-product matrix associated to $\mathbf{p}$.

%~ Static balance assumes zero acceleration thus we can write:
\begin{align} \label{eq:new_eul}
\underbrace{
\mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
\hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
}_\mathbf{G} \bm{\beta}, = 
\underbrace{\mat{\mathbf{0}_{3\times 3} \\ m \hat{\mathbf{g}}}}_{\mathbf{D}} \mathbf{c} + 
\underbrace{\mat{-m\mathbf{g} \\ \mathbf{0}}}_{\mathbf{d}}
\end{align}
where $\hat{\mathbf{x}} \in \R{3}{3}$ is the cross-product matrix associated to $\mathbf{x}$.
%~ The right hand side of \eqref{eq:new_eul} is independent of $\mathbf{c}^z$ because of the cross product matrix $\hat{\mathbf{g}}$ inside $\mathbf{D}$, so we can rewrite it as: \adnote{It is not really necessary to show that equilibrium is independent of the com altitude here, anyway the com is given, not a variable, so it does not change the size of the problem}
%~ \begin{align} \label{eq:new_eul2d}
%~ \mathbf{w}_0 = \mathbf{D}^{xy} \mathbf{c}^{xy} + \mathbf{d}
%~ \end{align}

@@

%~ In conclusion, the LP~\eqref{eq:dual} provides a method to check rapidly static equilibrium, and returns a robustness value.
%~ This robustness criterion can be immediately transformed into a heuristic, and eventually be weighted with others.
%~ Another option to enforce robustness is to set a minimum acceptable value for $b_0$, but as a result this conservative approach discards valid solutions.

>>>>

\section{A criterion for robust static equilibrium}
\label{sec:Equil}
%~ The planner is designed so that any generated contact configuration is in static equilibrium.
%~ We are interested in a robust 
%~ criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

We first give a linear program (LP) that verifies whether a contact configuration allows for static equilibrium. This LP is the same that was proposed in  \citep{Prete2016}.
From this formulation we derive a new LP that quantifies the robustness of the equilibrium to uncertainties in the contact forces.
In turn, from this value we can either choose the most robust candidate, or set a threshold on the required robustness. 
%~ While the presented LP is original, it is based on an analysis of the problem that we proposed in, where the interested reader can find more details.

\subsubsection{Conditions for static equilibrium}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
\item $m \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{\boldsymbol\upgamma}_{i1},\mathbf{\boldsymbol\upgamma}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{\boldsymbol\upgamma}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the non-slipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} -\mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i2} & \mathbf{n}_{i} - \mu \mathbf{\boldsymbol\upgamma}_{i2}}^T
\end{equation*}

Any force belonging to the linearized cone
can thus be expressed as a positive combination of its four generating rays.

@@

\begin{align} \label{eq:new_eul}
\underbrace{
\mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
\hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
}_\mathbf{G} \bm{\beta}, = 
\underbrace{\mat{\mathbf{0}_{3\times 3} \\ m \hat{\mathbf{g}}}}_{\mathbf{D}} \mathbf{c} + 
\underbrace{\mat{-m\mathbf{g} \\ \mathbf{0}}}_{\mathbf{d}}
\end{align}
where $\hat{\mathbf{x}} \in \R{3}{3}$ is the cross-product matrix associated to $\mathbf{x}$.

@@

### 2017-09-27 body/intro.tex
starting modifications

Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem.

The paper extends our previous papers presenting the reachability condition~\cite{tonneauisrr15} and the connection of our contact planner with a solver~$\mathcal{P}_3$~\cite{Carpentier2016}.
In this paper, we introduce a new efficient criterion of the robot balance, provide the complete pseudo-code and release our source code that can be used to directly reproduce our results.
We also provide a solution to $\mathcal{P}_3$.
The planner is validated on robot models of HyQ and HRP-2 (while \cite{tonneauisrr15} used unrealistic avatar models) with complete performance analysis.

>>>>

Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem. \\

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our conference paper~\cite{tonneauisrr15}. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section~\ref{rbprm} and~\ref{sec:contact} of the present paper are thus paraphrasing the original paper. However, we provide the pseudo-code of the algorithm,
absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation.
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.

To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work~\cite{Carpentier2016}. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in TODO.
We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand the experimental contribution of this paper.

Indeed, thanks to these tools we have been able to design a complete pipeline for synthesizing multi contact motions from scratch.
The novelty of the present paper lies in the rigorous experimental setup we propose to demonstrate its interest.
As opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are much more constrained kinematically than the virtual avatars used before,
thus making the planning problem even harder. Through an extensive empirical analysis, we demonstrate that the method remains interesting in this context, qualitatively and efficiently.
